How Consistent are LLMs in Applying Handcrafted Language Features

### 2024-06-11 article/main.tex
finish up first draft

For our experiment we will use state of the art large language models.
We used the Chatbot Arena~\cite{chiang2024chatbot} leaderboard for constructing
our selection.
In the interest of covering more ground, we choose large language models from
different vendors.
Our choice includes four top performing LLMs and one model that anyone would be
able to run on their own hardware.
The final selection is available in Table~\ref{table-llm}.

\begin{table}[ht]
    \setlength\tabcolsep{6pt}
    \centering
    \begin{tabular}{@{}lllll@{}}
        \toprule
        Name                    & Parameters  & Context Size & Elo Score & License Type        \\ \toprule
        OpenAI GPT-4o           & undisclosed & 128000       & 1287      & Proprietary         \\
        Google Gemini 1.5 Pro   & undisclosed & 1048576      & 1268      & Proprietary         \\
        Anthropic Claude 3 Opus & undisclosed & 200000       & 1248      & Proprietary         \\
        Meta Llama 3-8B         & 8 Billion   & 8000         & 1208      & Llama 3 Community   \\
        Cohere Command R+       & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
                                &             &              &           & with Acceptable Use \\
                                &             &              &           & Addendum            \\ \bottomrule
    \end{tabular}
    \caption{Experiment LLMs}\label{table-llm}
\end{table}

@@

In the next sections we only present the most relevant results of the
experiment.
To save space, we illustrate significant differences between the baseline and
the HLF instruction results in \textbf{bold} typeface.
The \underline{underlined} values represent situations when the HLF-enhanced
prompt lead to results that are closer to the corpus average than the baseline.
\textit{Italicized} values represent situations when the baseline was closer
to the target than the HLF-enhanced result.

@@

Table~\ref{table-cnn-dailymail-results} provide the most eloquent answer to this
question.

@@

            claude3 $Diff_N$    & \underline{\textbf{623.16}}  & \underline{\textbf{24.74}} & \underline{\textbf{51.18}}  & \underline{\textbf{26.68}} & 0.06                      & -0.0                      & \underline{\textbf{0.46}} & \underline{\textbf{1.4}}   & \underline{1.44}        & \textit{\textbf{-1.66}}   \\
            claude3 $Diff_A$    & \underline{\textbf{1774.0}}  & \underline{\textbf{74.0}}  & \underline{\textbf{158.84}} & \underline{\textbf{81.2}}  & 0.2                       & -0.01                     & \underline{\textbf{1.88}} & \underline{\textbf{4.49}}  & \underline{5.01}        & \textit{\textbf{-4.09}}   \\
            gemini $Diff_N$     & \underline{\textbf{985.86}}  & \underline{\textbf{34.52}} & \underline{\textbf{100.36}} & \underline{\textbf{53.39}} & \underline{\textbf{0.52}} & 0.05                      & \underline{\textbf{2.62}} & \underline{\textbf{4.68}}  & \textit{-4.17}          & \textit{\textbf{-2.43}}   \\
            gemini $Diff_A$     & \underline{\textbf{2990.5}}  & \underline{\textbf{106.5}} & \underline{\textbf{320.34}} & \underline{\textbf{159.1}} & \underline{\textbf{1.5}}  & 0.09                      & \underline{\textbf{7.24}} & \underline{\textbf{13.15}} & \textit{-12.65}         & \textit{\textbf{-7.13}}   \\
            gpt $Diff_N$        & \underline{\textbf{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textit{\textbf{-1.72}} & 0.43                      \\
            gpt $Diff_A$        & \underline{\textbf{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textit{\textbf{-5.72}} & 0.44                      \\
            llama3\_8b $Diff_N$ & 171.31                       & 6.63                       & -9.8                        & 13.8                       & 0.04                      & -0.04                     & 1.03                      & -0.84                      & -0.11                   & -0.76                     \\
            llama3\_8b $Diff_A$ & 524.0                        & 20.0                       & -32.5                       & 43.5                       & 0.1                       & -0.08                     & 3.64                      & -2.85                      & 1.57                    & -2.29                     \\
            command\_r $Diff_N$ & \underline{\textbf{1024.78}} & \underline{\textbf{39.67}} & \underline{\textbf{81.58}}  & \underline{\textbf{41.46}} & \underline{\textbf{0.46}} & \underline{\textbf{0.03}} & \underline{\textbf{2.67}} & \underline{\textbf{3.78}}  & \underline{1.4}         & \underline{\textbf{0.14}} \\
            command\_r $Diff_A$ & \underline{\textbf{3132.5}}  & \underline{\textbf{120.5}} & \underline{\textbf{246.84}} & \underline{\textbf{132.2}} & \underline{\textbf{1.36}} & \underline{\textbf{0.06}} & \underline{\textbf{7.88}} & \underline{\textbf{11.21}} & \underline{3.9}         & \underline{\textbf{1.03}} \\
            \bottomrule

@@

    }
    \caption{CNN-DailyMail Results}
    \label{table-cnn-dailymail-results}

@@

Command-R+ was the best performer by far, while Llama3 (8B) was predictibly the
worst at following HLF instructions.
Figure~\ref{fig-command-r-t-word} shows a positive result, where the result
obtained with the aid of HLF instructions is closer to the target.
Figure~\ref{fig-gpt-fkgl} shows a negative result, where the baseline is closer
to the target.
Figure~\ref{fig-llama-simp-ttr} shows an inconclusive result.

\begin{figure}[ht!]
    \centering
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail_t_word.png}
        \caption{Command-R+\\Total Words}
        \label{fig-command-r-t-word}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail_fkgl.png}
        \caption{GPT\\Flesch-Kincaid Grade Level}
        \label{fig-gpt-fkgl}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-llama3_8b-cnn_dailymail/prompt_1-llama3_8b-cnn_dailymail_simp_ttr.png}
        \caption[center]{Llama3-8B\\Simple TTR}
        \label{fig-llama-simp-ttr}
    \end{minipage}
\end{figure}

@@

In this variant of the experiment, we feed the LLM three examples from the
corpus.
\section{Discussion}

Discuss experiment results here.

@@

Yes, you can instruct at least some of the LLMs with HLFs.

compute BLEU, meteor, cider on styleptb using these LLMs and the prompt to
evaluate whether we actually achieve style transfer because of/in spite of HLF
instructions.
Human survey (amazon mechanical turk) on whether the text generations are good
or not. Likert scale (0 no good, 10 matches style perfectly)

>>>>

For our experiment we will use state of the art large language models.
We used the Chatbot Arena~\cite{chiang2024chatbot} leaderboard for constructing
our selection.
In the interest of covering more ground, we chose five top performers in the
Chatbot Arena from different vendors.
The final selection is available in Table~\ref{table-llm}.

\begin{table}[ht]
    \setlength\tabcolsep{6pt}
    \centering
    \begin{tabular}{@{}lllll@{}}
        \toprule
        Name                    & Parameters  & Context Size & Elo Score & License Type        \\ \toprule
        OpenAI GPT-4o           & undisclosed & 128000       & 1287      & Proprietary         \\
        Google Gemini 1.5 Pro   & undisclosed & 1048576      & 1268      & Proprietary         \\
        Anthropic Claude 3 Opus & undisclosed & 200000       & 1248      & Proprietary         \\
        Meta Llama 3-70B        & 70 Billion  & 8000         & 1208      & Llama 3 Community   \\
        Cohere Command R+       & 104 Billion & 128000       & 1189      & CC-BY-NC 4.0        \\
                                &             &              &           & with Acceptable Use \\
                                &             &              &           & Addendum            \\ \bottomrule
    \end{tabular}
    \caption{Large Language Model Selection}\label{table-llm}
\end{table}

@@

To save space, we illustrate significant differences between the baseline and
the HLF instruction results in \textbf{bold} typeface.
The \underline{underlined} values represent situations when the HLF-enhanced
prompt lead to results that are significantly closer to the corpus average than
the baseline.
\textit{Italicized} values represent situations when the baseline was
significantly closer to the target than the HLF-enhanced result.

In the next sections we only present the most relevant results of the
experiment.

@@

Table~\ref{table-prompt-1-cnn-dailymail} provide the most eloquent answer to
this question.

@@

            claude3 $Diff\_N$     & \textbf{\underline{623.16}}  & \textbf{\underline{24.74}} & \textbf{\underline{51.18}}  & \textbf{\underline{26.68}} & 0.06                      & -0.0                      & \textbf{\underline{0.46}} & \textbf{\underline{1.4}}   & 1.44                    & \textbf{\textit{-1.66}}   \\ \midrule
            claude3 $Diff\_A$     & \textbf{\underline{1774.0}}  & \textbf{\underline{74.0}}  & \textbf{\underline{158.84}} & \textbf{\underline{81.2}}  & 0.2                       & -0.01                     & \textbf{\underline{1.88}} & \textbf{\underline{4.49}}  & 5.01                    & \textbf{\textit{-4.09}}   \\ \midrule
            gemini $Diff\_N$      & \textbf{\underline{985.86}}  & \textbf{\underline{34.52}} & \textbf{\underline{100.36}} & \textbf{\underline{53.39}} & \textbf{\underline{0.52}} & 0.05                      & \textbf{\underline{2.62}} & \textbf{\underline{4.68}}  & -4.17                   & \textbf{\textit{-2.43}}   \\ \midrule
            gemini $Diff\_A$      & \textbf{\underline{2990.5}}  & \textbf{\underline{106.5}} & \textbf{\underline{320.34}} & \textbf{\underline{159.1}} & \textbf{\underline{1.5}}  & 0.09                      & \textbf{\underline{7.24}} & \textbf{\underline{13.15}} & -12.65                  & \textbf{\textit{-7.13}}   \\ \midrule
            gpt $Diff\_N$         & \textbf{\underline{125.87}}  & 1.46                       & 10.95                       & 10.77                      & 0.02                      & -0.01                     & 0.71                      & 0.81                       & \textbf{\textit{-1.72}} & 0.43                      \\ \midrule
            gpt $Diff\_A$         & \textbf{\underline{327.0}}   & 3.0                        & 29.0                        & 27.0                       & 0.08                      & -0.01                     & 1.82                      & 2.24                       & \textbf{\textit{-5.72}} & 0.44                      \\ \midrule
            command\_r $Diff\_N$  & \textbf{\underline{1024.78}} & \textbf{\underline{39.67}} & \textbf{\underline{81.58}}  & \textbf{\underline{41.46}} & \textbf{\underline{0.46}} & \textbf{\underline{0.03}} & \textbf{\underline{2.67}} & \textbf{\underline{3.78}}  & 1.4                     & \textbf{\underline{0.14}} \\ \midrule
            command\_r $Diff\_A$  & \textbf{\underline{3132.5}}  & \textbf{\underline{120.5}} & \textbf{\underline{246.84}} & \textbf{\underline{132.2}} & \textbf{\underline{1.36}} & \textbf{\underline{0.06}} & \textbf{\underline{7.88}} & \textbf{\underline{11.21}} & 3.9                     & \textbf{\underline{1.03}} \\ \midrule
            llama3\_70b $Diff\_N$ & \textbf{\underline{459.45}}  & \textbf{\underline{17.78}} & \textbf{\underline{36.5}}   & \textbf{\underline{19.75}} & \textbf{\underline{0.12}} & \textbf{\underline{0.03}} & \textbf{\underline{1.43}} & \textbf{\underline{2.18}}  & 1.5                     & \textbf{\textit{-0.06}}   \\ \midrule
            llama3\_70b $Diff\_A$ & \textbf{\underline{1363.5}}  & \textbf{\underline{54.0}}  & \textbf{\underline{106.5}}  & \textbf{\underline{60.5}}  & \textbf{\underline{0.36}} & \textbf{\underline{0.09}} & \textbf{\underline{4.33}} & \textbf{\underline{6.38}}  & 7.46                    & \textbf{\textit{-0.11}}   \\ \bottomrule

@@

    }\caption{CNN Results}\label{table-prompt-1-cnn-dailymail}

@@

For our selection of features, on the CNN corpus of articles, Command-R+ was the
best performer.
GPT scored surprisingly low.

Figure~\ref{fig-command-r-t-word} shows a positive result, where the result
obtained with the aid of HLF instructions is closer to the target.
Figure~\ref{fig-gpt-fkgl} shows a negative result, where the baseline is closer
to the target.
Figure~\ref{fig-llama-simp-ttr} shows an inconclusive result.

\begin{figure}[ht!]
    \centering
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-command_r-cnn_dailymail/prompt_1-command_r-cnn_dailymail_t_word.png}
        \caption{Command-R+\\Total Words}
        \label{fig-command-r-t-word}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail_fkgl.png}
        \caption{GPT\\Flesch-Kincaid Grade Level}
        \label{fig-gpt-fkgl}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail_a_verb_pw.png}
        \caption[center]{Claude3\\Average Verbs per Word}
        \label{fig-llama-simp-ttr}
    \end{minipage}
\end{figure}

@@

The second type of instructional prompt we feed to the LLM contains three
examples from the corpus.
The hypothesis is that this will influence the baseline result to be closer to
the corpus average.

In this case, it is interesting to see how the LLM generates text based on input
text from both within the corpus that contains the example articles and from
outside the corpus.

In Table~\ref{table-prompt-2-cnn-dailymail} we see the results when an input from
outside the corpus is used.
The outlook is similar for what we saw in the first experiment with GPT
being the least responsive to HLF instructions.

\begin{table}[!ht]
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{lllllllllll}
            model                 & t\_word                      & t\_sent                    & n\_uverb                    & n\_uadj                    & simp\_ttr                 & a\_verb\_pw               & corr\_adj\_var            & corr\_verb\_var           & fkgl                       & a\_kup\_pw                \\
            \toprule
            claude3 $Diff\_N$     & \textbf{\underline{718.49}}  & \textbf{\underline{24.13}} & \textbf{\underline{44.43}}  & \textbf{\underline{23.43}} & \textbf{\underline{0.4}}  & 0.03                      & 0.94                      & \textbf{\underline{0.77}} & 1.2                        & 0.07                      \\ \midrule
            claude3 $Diff\_A$     & \textbf{\underline{2345.54}} & \textbf{\underline{90.18}} & \textbf{\underline{148.36}} & \textbf{\underline{63.6}}  & \textbf{\underline{1.18}} & 0.07                      & 1.36                      & \textbf{\underline{2.55}} & 3.33                       & 0.52                      \\ \midrule
            gemini $Diff\_N$      & \textbf{\underline{785.6}}   & \textbf{\underline{25.8}}  & \textbf{\underline{60.5}}   & \textbf{\underline{41.6}}  & \textbf{\underline{0.41}} & \textbf{\underline{0.05}} & \textbf{\underline{1.95}} & \textbf{\underline{2.67}} & -4.62                      & \textbf{\textit{-1.94}}   \\ \midrule
            gemini $Diff\_A$      & \textbf{\underline{2512.0}}  & \textbf{\underline{81.5}}  & \textbf{\underline{210.34}} & \textbf{\underline{125.7}} & \textbf{\underline{1.25}} & \textbf{\underline{0.09}} & \textbf{\underline{5.84}} & \textbf{\underline{8.05}} & -9.46                      & \textbf{\textit{-5.82}}   \\ \midrule
            gpt $Diff\_N$         & 47.8                         & 5.18                       & 6.13                        & -0.13                      & 0.02                      & 0.0                       & 0.05                      & 0.3                       & 3.06                       & -0.27                     \\ \midrule
            gpt $Diff\_A$         & 77.0                         & 12.5                       & 13.0                        & -4.0                       & 0.04                      & 0.0                       & -0.17                     & 0.77                      & 8.45                       & -0.16                     \\ \midrule
            command\_r $Diff\_N$  & \textbf{\underline{263.51}}  & 5.35                       & \textbf{\underline{25.08}}  & \textbf{\underline{17.37}} & \textbf{\textit{-0.09}}   & -0.01                     & \textbf{\underline{1.74}} & \textbf{\underline{1.42}} & \textbf{\underline{10.46}} & \textbf{\underline{1.56}} \\ \midrule
            command\_r $Diff\_A$  & \textbf{\underline{896.0}}   & 20.5                       & \textbf{\underline{86.0}}   & \textbf{\underline{60.6}}  & \textbf{\textit{-0.21}}   & -0.03                     & \textbf{\underline{5.45}} & \textbf{\underline{4.95}} & \textbf{\underline{31.29}} & \textbf{\underline{5.67}} \\ \midrule
            llama3\_70b $Diff\_N$ & \textbf{\underline{376.0}}   & \textbf{\underline{15.63}} & \textbf{\underline{38.63}}  & 15.74                      & \textbf{\underline{0.24}} & \textbf{\underline{0.06}} & 1.69                      & \textbf{\underline{2.91}} & 0.31                       & -0.3                      \\ \midrule
            llama3\_70b $Diff\_A$ & \textbf{\underline{1088.5}}  & \textbf{\underline{45.0}}  & \textbf{\underline{113.0}}  & 41.0                       & \textbf{\underline{0.65}} & \textbf{\underline{0.18}} & 3.53                      & \textbf{\underline{8.56}} & 0.27                       & -0.68                     \\ \bottomrule
        \end{tabular}%
    }
    \caption{prompt 2 - CNN Results\\Input Outside Corpus}
    \label{table-prompt-2-cnn-dailymail}
\end{table}

GPT's results would be good if the baseline were very close to the target value.
As we can see in Figure~\ref{fig-p2-gpt-twords}, this is not always the case.
We also notice that Command-R+(fig.~\ref{fig-p2-command-r-simpttr}) and Gemini
(fig.~\ref{fig-p2-gemini-a-kup-pw}) both registered significantly negative
results, in contrast to their performance in our first scenario.

\begin{figure}[ht!]
    \centering
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail_t_word.png}
        \caption{GPT\\Total Words}
        \label{fig-p2-gpt-twords}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-command_r-cnn_dailymail/prompt_2-command_r-cnn_dailymail_simp_ttr.png}
        \caption{Command-R+\\Simple TTR}
        \label{fig-p2-command-r-simpttr}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_a_kup_pw.png}
        \caption[center]{Gemini\\Kuperman Age of Aquisition}
        \label{fig-p2-gemini-a-kup-pw}
    \end{minipage}
\end{figure}

The results obtained in our first scenario are largely confirmed by our second
scenario.
So now we turn to see how using a text from within the corpus affects the
performance of the large language models.
The hypothesis is that if the text is `familiar', both the baseline and the HLF
instruction prompt will produce results that are much closer to the target
value.
The expectation is to see less significant results.

\begin{table}[!ht]
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{lllllllllll}
            model                 & t\_word                      & t\_sent                     & n\_uverb                    & n\_uadj                    & simp\_ttr                 & a\_verb\_pw               & corr\_adj\_var             & corr\_verb\_var            & fkgl                       & a\_kup\_pw              \\
            \toprule
            claude3 $Diff\_N$     & \textbf{\underline{871.02}}  & \textbf{\underline{34.2}}   & \textbf{\underline{34.48}}  & \textbf{\textit{-7.38}}    & \textbf{\underline{0.27}} & 0.01                      & \textbf{\textit{-0.93}}    & \textbf{\textit{-0.72}}    & 1.16                       & -0.14                   \\ \midrule
            claude3 $Diff\_A$     & \textbf{\underline{2756.12}} & \textbf{\underline{117.18}} & \textbf{\underline{126.36}} & \textbf{\textit{-11.2}}    & \textbf{\underline{0.81}} & -0.01                     & \textbf{\textit{-2.52}}    & \textbf{\textit{-0.74}}    & 3.69                       & -0.4                    \\ \midrule
            gemini $Diff\_N$      & \textbf{\underline{1315.41}} & \textbf{\underline{46.19}}  & \textbf{\underline{114.47}} & \textbf{\underline{72.38}} & \textbf{\underline{1.17}} & 0.03                      & \textbf{\underline{5.25}}  & \textbf{\underline{6.64}}  & -1.6                       & \textbf{\textit{-2.1}}  \\ \midrule
            gemini $Diff\_A$      & \textbf{\underline{3996.0}}  & \textbf{\underline{138.0}}  & \textbf{\underline{357.36}} & \textbf{\underline{212.7}} & \textbf{\underline{3.3}}  & 0.06                      & \textbf{\underline{15.49}} & \textbf{\underline{19.4}}  & -13.22                     & \textbf{\textit{-6.19}} \\ \midrule
            gpt $Diff\_N$         & \textbf{\underline{521.29}}  & \textbf{\underline{23.3}}   & \textbf{\underline{43.13}}  & 19.85                      & \textbf{\underline{0.2}}  & 0.02                      & 1.08                       & \textbf{\underline{1.04}}  & 5.09                       & 0.67                    \\ \midrule
            gpt $Diff\_A$         & \textbf{\underline{1464.0}}  & \textbf{\underline{65.5}}   & \textbf{\underline{121.84}} & 58.1                       & \textbf{\underline{0.58}} & 0.05                      & 2.93                       & \textbf{\underline{2.59}}  & 14.23                      & 1.89                    \\ \midrule
            command\_r $Diff\_N$  & \textbf{\underline{371.08}}  & \textbf{\underline{13.31}}  & 22.06                       & \textbf{\underline{17.19}} & \textbf{\underline{0.17}} & \textbf{\underline{0.02}} & 0.88                       & 1.09                       & \textbf{\underline{4.05}}  & -0.15                   \\ \midrule
            command\_r $Diff\_A$  & \textbf{\underline{1198.0}}  & \textbf{\underline{42.5}}   & 73.0                        & \textbf{\underline{56.5}}  & \textbf{\underline{0.69}} & \textbf{\underline{0.07}} & 2.85                       & 3.99                       & \textbf{\underline{13.89}} & -0.11                   \\ \midrule
            llama3\_70b $Diff\_N$ & \textbf{\underline{1040.85}} & \textbf{\underline{41.99}}  & \textbf{\underline{80.58}}  & \textbf{\underline{49.12}} & \textbf{\underline{0.55}} & \textbf{\underline{0.04}} & \textbf{\underline{3.99}}  & \textbf{\underline{5.05}}  & 7.18                       & \textbf{\textit{-1.94}} \\ \midrule
            llama3\_70b $Diff\_A$ & \textbf{\underline{2928.0}}  & \textbf{\underline{118.5}}  & \textbf{\underline{222.0}}  & \textbf{\underline{133.5}} & \textbf{\underline{1.43}} & \textbf{\underline{0.05}} & \textbf{\underline{10.77}} & \textbf{\underline{13.42}} & 17.21                      & \textbf{\textit{-6.2}}  \\ \bottomrule
        \end{tabular}%
    }
    \caption{CNN Results\\Input Inside Corpus}
    \label{table-prompt-2-ifd-cnn-dailymail}
\end{table}

The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
expectations outright.
The most surprising behaviour is exhibited in the context of more concrete HLFs
such as the total number of words or the number of unique adjectives.
Claude3 shows regressive behaviour for the number of unique adjectives in
Figure~\ref{fig-p2-ifd-claude3-nuadj}, while Gemini constructs texts using a
very different amount of words in Figure~\ref{fig-p2-ifd-gemini-twords}.
Less surprisingly, Llama3 fails to generate text that conforms to the correct
Kuperman age of acquisition per word in Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.

\begin{figure}[ht!]
    \centering
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail_n_uadj.png}
        \caption{Claude3\\Number of Unique Adjectives}
        \label{fig-p2-ifd-claude3-nuadj}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_t_word.png}
        \caption{Gemini\\Total Words}
        \label{fig-p2-ifd-gemini-twords}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
        \caption[center]{Llama3\\Kuperman Age of Aquisition}
        \label{fig-p2-ifd-llama3-a-kup-pw}
    \end{minipage}
\end{figure}

The results in this last scenario show that whether LLMs take examples listed in
the prompt into account or not with respect to conforming to certain linguistic
features is debatable, providing them with specific instructions derived from
HLFs makes a significant impact.

\section{Discussion}

In terms of our first question --- whether LLMs are able to follow instructions
concerning handcrafted linguistic features --- the response is in our opinion
affirmative.
Certainly, not all LLMs are created equal in this regard.
It is debatable whether this experiment is well designed because of the choice
to derive target HLF values for the LLM from the corpus.
This is evident in the results obtained for the Yelp dataset.
While these results also show a significant influence when using HLFs in the
prompt, the influence is mostly negative.
This can be attributed to the choice of input text in relation to the corpus,
the wording in the instruction prompt or to each model's specific training.

This leads us to our second question: how much do handcrafted linguistic feature
instructions affect the way LLMs write text.
The results in this regards show us a much less conclusive picture.

Firstly, we cannot assume that the LLM will follow HLF instructions the way we
intend.
Whether we obtain the expected outcome or not depends largely on the prompt,
the chosen HLFs, the model, the target values for the HLFs, a.s.o
We do, however, notice a few patterns.
The more abstract the instruction, the less it seems to make a positve impact.
More importantly, by inspecting the result data, a more abstract instruction
seems more likely to produce a negative impact on the desired outcome.

Secondly, the choice of input text in relation to the target HLF values seems
very important.
Choosing to reword a poem so that it has the same reading ease as a user review
seems like a setup for poor results.
This leads us to intuit that LLMs are clearly able to use HLF instructions to
generate text differently, but there is a limit to their `creativity'.

Finally, the choice of HLFs is very important.
The text generated by an LLM is less likely to conform to HLFs that result in
more abstract instructions.
Furthermore, the combination of chosen HLFs is important.
Choosing HLFs that measure the uniqueness of a linguistic element (words,
structures, a.s.o) limits the LLMs ability to conform to other HLFs (for example
total number of sentences or reading ease).
This is highlighted in the experiment specifically in the tension between the
age of acquisition and the adjective or verb variation.
Predictably, the more concrete instructions take precedence over the more
abstract ones.

@@

We started out by trying to understand whether it is possible to conform the
output of a large language model to exhibit specific linguistic features.
To our surprise, it turns out that instructing language models to write text
that conforms to a selected assortment of linguistic features almost always
yields significant changes in how the text is being generated.

This doesn't mean that the results are always satisfying.
What the model writes about and the chosen target values for the linguistic
features are major factors in how successful the overall experience is.
Unfortunately, at this stage, finding out good target values and good writing
instructions is a question of trial and error.

Other experiment designs might have been more opportune in examining the
questions layed out in the introduction.
For example, using set target values for HLFs (without deriving them from a
corpus of text) and choosing input text where the HLFs vary greatly would have
been an alternate design.
Following through on those approaches is one of the possible expansions on the
current work.

Another way of expanding on the current work is finding the settings in which
large language models are most responsive to linguistic features.
The current experiment was inconclusive in this regard and a legitimate further
question is whether settings in which text generated by language models
consistently conforms to certain linguistic features actually exist.

Lastly, a simpler question that is left unanswered relates to the number of
linguistic features that were used.
What if we would have measured models' behaviour for single linguistic features
independently?
